---
title: "Read Txdot Collision Data"
author: "Alan Jackson"
date: "February 26, 2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)

googlecrs <- 4326

#   Crash Severity ID table
Severity <- c(
"0",	"Unknown", 
"1",	"Suspected Serious Injury",
"2",	"Non-Incapacitating",
"3",	"Possible Injury",
"4",	"Fatal",
"5",	"Not Injured",
"94",	"Reported Invalid",
"95",	"Not Reported"
)
Severity <- cbind.data.frame(split(Severity, rep(1:2, times=length(Severity)/2)), stringsAsFactors=F)
names(Severity) <- c("Crash_Sev_ID", "Severity")
Severity$Crash_Sev_ID <- as.integer(Severity$Crash_Sev_ID)

#   Intersection ID table
Intersection <- c(
"1",	"Intersection", 
"2",	"Intersection Related",
"3",	"Driveway Access",
"4",	"Non-Intersection",
"5",	"Not Reported",
"93",	"Undetermined",
"94",	"Reported Invalid"
)
Intersection <- cbind.data.frame(split(Intersection, rep(1:2, times=length(Intersection)/2)), stringsAsFactors=F)
names(Intersection) <- c("Intrsct_Relat_ID", "Intersection")
Intersection$Intrsct_Relat_ID <- as.integer(Intersection$Intrsct_Relat_ID)


#   Road Type ID table
RoadType <- c(
  "1", "Interstate",
  "2", "US & State Highways",
  "3", "Farm To Market",
  "4", "County Road",
  "5", "City Street",
  "6", "Tollway",
  "7", "Other Roads",
  "8", "Toll Bridges",
  "9", "Non Trafficway",
  "93", "Undetermined - Failed Business Rule(s)",
  "94", "Reported Invalid",
  "95", "Not Reported"
)
RoadType <- cbind.data.frame(split(RoadType, rep(1:2, times=length(RoadType)/2)), stringsAsFactors=F)
names(RoadType) <- c("Road_Cls_ID", "RoadType")
RoadType$Road_Cls_ID <- as.integer(RoadType$Road_Cls_ID)


#   Crash Harm ID table
Harm <- c(
"1",	"Pedestrian", 
"2",	"Motor Vehicle in Transport",
"3",	"Rail Road",
"4",	"Parked Car",
"5",	"Pedal Cyclist",
"6",	"Animal",
"7",	"Fixed Object",
"8",	"Other Object",
"9",	"Other Non-Collision",
"10",	"Overturned",
"11",	"Not Reported",
"93",	"Undetermined",
"94",	"Reported Invalid"
)
Harm <- cbind.data.frame(split(Harm, rep(1:2, times=length(Harm)/2)), stringsAsFactors=F)
names(Harm) <- c("Harm_Evnt_ID", "Harmed")
Harm$Harm_Evnt_ID <- as.integer(Harm$Harm_Evnt_ID)

#   Charge Category ID table
Charge <- c(
"10", "ALCOHOL/DRUGS",
"11", "ASSAULT/MANSLAUGHTER WITH VEHICLE",
"12",	"BICYCLE OFFENSES",
"13",	"COMMERCIAL & MOTOR CARRIER VIOLATIONS",
"14",	"DRIVER LICENSE & REGISTRATION VIOLATIONS",
"15",	"DRIVER'S MISCELLANEOUS VIOLATIONS",
"16",	"INSURANCE AND INSPECTION",
"17",	"MOTORCYCLE VIOLATIONS",
"18",	"OVERTAKING (PASSING)",
"19",	"PARKING VIOLATIONS",
"20",	"REPORTING VIOLATIONS (TRAFFIC VIOLATIONS)",
"21",	"RIGHT OF WAY",
"22",	"SEATBELT / HELMET USAGE",
"23",	"SIGNAL INTENTION",
"24",	"SPEEDING",
"25",	"TRAFFIC SIGNS, SIGNALS, AND ROAD MARKINGS",
"26",	"TURNING VIOLATIONS",
"27",	"VEHICLE/EQUIPMENT DEFECTS",
"28",	"WRONG SIDE/WRONG WAY",
"29",	"OTHER",
"30",	"NONE",
"93",	"UNDETERMINED - FAIL BR",
"94",	"REPORTED INVALID",
"95",	"NOT REPORTED"
)
Charge <- cbind.data.frame(split(Charge, rep(1:2, times=length(Charge)/2)), stringsAsFactors=F)
names(Charge) <- c("Charge_Cat_ID", "Charge")
Charge$Charge_Cat_ID <- as.integer(Charge$Charge_Cat_ID)




HarrisBoundary <- c(-95.990842, 29.499559, -94.852384, 30.180927)

City <- c(
 "29", "Baytown",
 "30", "Beaumont",
 "33", "Bellaire",
 "57", "Bunker Hill Village",
 "111", "Deer Park",
 "135", "El Lago",
 "160", "Friendswood",
 "164", "Galena Park",
 "208", "Houston",
 "211", "Humble",
 "212", "Hunters Creek Village",
 "219", "Jacinto City",
 "223", "Jersey Village",
 "228", "Katy",
"244", "La Porte",
"255", "League City",
"260", "Liberty",
"296", "Missouri City",
"304", "Nassau Bay",
"325", "Pasadena",
"326", "Pearland",
"333", "Piney Point Village",
"359", "Richmond",
"383", "San Marcos",
"391", "Seabrook",
"405", "South Houston",
"408", "Spring Valley",
"409", "Stafford",
"422", "Taylor Lake Village",
"430", "Tomball",
"434", "Tyler",
"447", "Webster",
"453", "West University Place",
"478", "Aldine",
"625", "Channelview",
"647", "Cloverleaf",
"672", "Crosby",
"843", "Hedwig Village",
"851", "Highlands",
"855", "Hilshire Village",
"1061", "Mission Bend",
"1073", "Morgans Point",
"1326", "Shoreacres",
"1348", "Southside Place",
"1353", "Spring",
"1444", "Waller",
"1535", "Kingwood",
"1585", "Rural Brazoria County",
"1601", "Rural Chambers County",
"1644", "Rural Fort Bend County",
"1649", "Rural Galveston County",
"1665", "Rural Hardin County",
"1666", "Rural Harris County",
"1667", "Rural Harrison County",
"1668", "Rural Hartley County",
"1678", "Rural Houston County",
"1711", "Rural Liberty County",
"1719", "Rural Madison County",
"1735", "Rural Montgomery County",
"1926", "Alief",
"2071", "Atascosita",
"3313", "Cypress",
"4728", "Huffman",
"5659", "Meadows Place",
"6756", "Porter"
)
City <- cbind.data.frame(split(City, rep(1:2, times=length(City)/2)), stringsAsFactors=F)
names(City) <- c("Rpt_City_ID", "City")
City$Rpt_City_ID <- as.integer(City$Rpt_City_ID)

knitr::opts_chunk$set(echo = TRUE)
```

##  TXDoT Crash files

Database is documented at https://www.txdot.gov/inside-txdot/division/traffic/data-access.html

Access is from https://cris.txdot.gov/secure/Share

Log on and download one year at a time

After unzipping, main files required for crash data are the files with "crash" in the filename. Also need to add the data from the "charges" files.

The "unit" files contain information about the car (make, model, year), which I will ignore for now.

Download files and unzip all of them into a single directory.




```{r Read files}
filenames <- list.files(path = "~/Dropbox/Rprojects/Collisions/TrafficAccidents/fullunzip", pattern="*csv$")

filenames <- paste("~/Dropbox/Rprojects/Collisions/TrafficAccidents/fullunzip/",filenames, sep="")

#filenames <- filenames[grep("201[01]0101", filenames)] # for testing

#  first read in the crash files

DF_collisions <- filenames[grep("crash", filenames)] %>% 
  map(~read_csv(.x, col_types = cols(
    Crash_Date = col_date(format="%m/%d/%Y"),
    Crash_Time = col_time(format = "%H:%M %p")
  ))) %>% 
  reduce(rbind)

#  now read in the charges files

DF_charges <- filenames[grep("charges", filenames)] %>% 
  map(read_csv) %>% 
  reduce(rbind)

#   The Charge_Cat_ID is all but useless, as it is not filled in a
#   majority of the time when Charge is filled in. So we will need
#   to do some work.

```

##  Analyze charges file and simplify it so we can use it.


```{r Look at Charges file}

ChargesTemp <- DF_Charges_temp[DF_Charges_temp$Simple_Charge=="",] %>% group_by(Charge) %>% tally() %>% arrange(-n)

# Start a new field, Simple_Charge which will be used
# But first populate a:z with each charge, and combine later

DF_charges$a[DF_charges$Charge=="NO CHARGES"] <- "None"
DF_charges$a[DF_charges$Charge=="NO CHARGE"] <- "None"
DF_charges$a[DF_charges$Charge=="NONE"] <- "None"
DF_charges$a[DF_charges$Charge=="NONE."] <- "None"
DF_charges$a[DF_charges$Charge=="NONE FILED"] <- "None"
DF_charges$a[DF_charges$Charge=="NONE AT THIS TIME"] <- "None"
DF_charges$a[DF_charges$Charge=="N/A"] <- "None"
DF_charges$a[grepl("CONFLICT[IE][ND]G? STATEMENT", DF_charges$Charge)] <- "None"
DF_charges$a[grepl("NOT? CIT", DF_charges$Charge)] <- "None"
DF_charges$a[grepl("NO CHARGES", DF_charges$Charge)] <- "None"
DF_charges$a[grepl("NOT? CH", DF_charges$Charge)] <- "None"
DF_charges$a[grepl("^ACCIDENT$", DF_charges$Charge)] <- "None"
DF_charges$a[grepl("^CRASH$", DF_charges$Charge)] <- "None"
DF_charges$a[grepl("^ACC$", DF_charges$Charge)] <- "None"
DF_charges$a[is.na(DF_charges$Charge)] <- "None"


#   Function to build mask for updating Simple_Charge
update_charge <- function(match){
  mask <- grepl(match, DF_charges$Charge)
  print(sum(mask))
  mask
}
# Sometimes there are multiple violations squeezed into the field. This
# is a problem. I will pick based on which is most common, e.g., 
# speeding beats no license, and work it that way.

# Speeding
mask <- update_charge("SPEE*D")
DF_charges$b[mask] <- "Speeding"

mask <- update_charge("SPEE ")
DF_charges$b[mask] <- "Speeding"

mask <- update_charge("SPEE")
DF_charges$b[mask] <- "Speeding"

mask <- update_charge("SPD")
DF_charges$b[mask] <- "Speeding"

mask <- update_charge("SEED")
DF_charges$b[mask] <- "Speeding"

mask <- update_charge("SPPED")
DF_charges$b[mask] <- "Speeding"

mask <- update_charge("F\\.?T\\.?C\\.?S")
DF_charges$b[mask] <- "Speeding"

# Bad license plates (do this first so can use license to find drivers license)
mask <- update_charge("PLATE")
DF_charges$c[mask] <- "Bad Plates"

# No or expired driver's license
mask <- update_charge("LICENSE")
DF_charges$d[mask] <- "No License"

mask <- update_charge(" DL")
DF_charges$d[mask] <- "No License"

mask <- update_charge(" TDL")
DF_charges$d[mask] <- "No License"

mask <- update_charge(" LIC ")
DF_charges$d[mask] <- "No License"

mask <- update_charge("LICENC")
DF_charges$d[mask] <- "No License"

mask <- update_charge("LISCEN")
DF_charges$d[mask] <- "No License"

mask <- update_charge("LINCEN")
DF_charges$d[mask] <- "No License"

mask <- update_charge("TXDL")
DF_charges$d[mask] <- "No License"

mask <- update_charge("NO OPER")
DF_charges$d[mask] <- "No License"

mask <- update_charge("D\\.L")
DF_charges$d[mask] <- "No License"

mask <- update_charge("D\\.?W\\.?L")
DF_charges$d[mask] <- "No License"

mask <- update_charge("NDL")
DF_charges$d[mask] <- "No License"

mask <- update_charge("NO O\\.?P?L")
DF_charges$d[mask] <- "No License"

mask <- update_charge("N\\.?O\\.?L")
DF_charges$d[mask] <- "No License"

mask <- update_charge("CDL")
DF_charges$d[mask] <- "No License"

mask <- update_charge("^NO L$")
DF_charges$d[mask] <- "No License"

mask <- update_charge(" D/L")
DF_charges$d[mask] <- "No License"

mask <- update_charge("^DL ")
DF_charges$d[mask] <- "No License"

mask <- update_charge("^N?O?T?/?-?DL$")
DF_charges$d[mask] <- "No License"

mask <- update_charge("DWLI")
DF_charges$d[mask] <- "No License"

mask <- update_charge("ENDOR")
DF_charges$d[mask] <- "No License"

mask <- update_charge("LISEN")
DF_charges$d[mask] <- "No License"

mask <- update_charge("UNLI")
DF_charges$d[mask] <- "No License"

mask <- update_charge("DRIVER'?S? LIC")
DF_charges$d[mask] <- "No License"

# No insurance
mask <- update_charge("INSUR")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("^INS$")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("INSR")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("^NO[ -/]INS")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("RESP")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("NO F\\.R\\.")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("FINAN")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("FIN RES")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("MAINTAIN F")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("NO INS\\.?$")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("NO INS\\.")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("NO INSU$")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("FT?MFR")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("F\\.?M")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("F\\.T\\.?M")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("LIAB")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("F M F R")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("QUALI")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("NOIN?$")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("NOIN?$")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("PROOF OF FI")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge(" NSURA")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("IN?SUR?ANCE")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("URANCE")
DF_charges$e[mask] <- "No Insurance"

mask <- update_charge("PROOF OF INS")
DF_charges$e[mask] <- "No Insurance"

# Ran red light
mask <- update_charge(" RED ")
DF_charges$f[mask] <- "Ran Red Light"

mask <- update_charge("RED ?LIGHT")
DF_charges$f[mask] <- "Ran Red Light"

mask <- update_charge(" RED$")
DF_charges$f[mask] <- "Ran Red Light"

mask <- update_charge("STOP LIGHT")
DF_charges$f[mask] <- "Ran Red Light"

mask <- update_charge("CONTROL DEV")
DF_charges$f[mask] <- "Ran Red Light"

mask <- update_charge("STOP ?AND|\\&|/ ?GO")
DF_charges$f[mask] <- "Ran Red Light"

mask <- update_charge("LANE CONT")
DF_charges$f[mask] <- "Ran Red Light"

mask <- update_charge("DISREGARDE?D? SIGNAL")
DF_charges$f[mask] <- "Ran Red Light"

mask <- update_charge("DISREGARDE?D? LIGHT")
DF_charges$f[mask] <- "Ran Red Light"

mask <- update_charge("DISREGARDE?D? TRAFFIC")
DF_charges$f[mask] <- "Ran Red Light"

mask <- update_charge("RAN A? ?LIGHT")
DF_charges$f[mask] <- "Ran Red Light"

mask <- update_charge("RAN A? ?SIGNAL LIGHT")
DF_charges$f[mask] <- "Ran Red Light"

mask <- update_charge("RAN RA?D ")
DF_charges$f[mask] <- "Ran Red Light"

# Unsafe Lane Change
mask <- update_charge("CHANGES? LANE")
DF_charges$g[mask] <- "Unsafe Lane Change"

mask <- update_charge("CHANGED LANE")
DF_charges$g[mask] <- "Unsafe Lane Change"

mask <- update_charge("LANE CHANGE")
DF_charges$g[mask] <- "Unsafe Lane Change"

mask <- update_charge("CHANGING LANE")
DF_charges$g[mask] <- "Unsafe Lane Change"

mask <- update_charge("CHANGED? ?LA?NE?")
DF_charges$g[mask] <- "Unsafe Lane Change"

mask <- update_charge("LA?NE? ?CHANGE")
DF_charges$g[mask] <- "Unsafe Lane Change"

# Single Lane failure
mask <- update_charge("SI[NG][GN]LE")
DF_charges$h[mask] <- "Fail Single Lane"

mask <- update_charge("SINGL ")
DF_charges$h[mask] <- "Fail Single Lane"

mask <- update_charge("MAINTAIN SING ")
DF_charges$h[mask] <- "Fail Single Lane"

mask <- update_charge("MAINTAIN LANE")
DF_charges$h[mask] <- "Fail Single Lane"

# DWI
mask <- update_charge("D\\.?W\\.?I")
DF_charges$i[mask] <- "DWI"

mask <- update_charge("INTOX")
DF_charges$i[mask] <- "DWI"

mask <- update_charge("D\\.?U\\.?I")
DF_charges$i[mask] <- "DWI"

mask <- update_charge("INFLU")
DF_charges$i[mask] <- "DWI"

# Failure to Yield
mask <- update_charge("YIELD")
DF_charges$j[mask] <- "Failure to Yield"

mask <- update_charge("YLD")
DF_charges$j[mask] <- "Failure to Yield"

mask <- update_charge("YL ")
DF_charges$j[mask] <- "Failure to Yield"

mask <- update_charge("YEIL")
DF_charges$j[mask] <- "Failure to Yield"

mask <- update_charge("YIEL")
DF_charges$j[mask] <- "Failure to Yield"

mask <- update_charge("FYR")
DF_charges$j[mask] <- "Failure to Yield"

mask <- update_charge("FTYR")
DF_charges$j[mask] <- "Failure to Yield"

mask <- update_charge("^F\\.?T\\.?Y ")
DF_charges$j[mask] <- "Failure to Yield"

mask <- update_charge("FAIL TO Y")
DF_charges$j[mask] <- "Failure to Yield"

mask <- update_charge("FAIL TO STOP FR")
DF_charges$j[mask] <- "Failure to Yield"

mask <- update_charge("YROW")
DF_charges$j[mask] <- "Failure to Yield"

# Failure to Stop and Give Information
mask <- update_charge("F\\.?S\\.?G\\.?I")
DF_charges$k[mask] <- "FSGI"

mask <- update_charge("GIVE INFO")
DF_charges$k[mask] <- "FSGI"

mask <- update_charge("FGSI")
DF_charges$k[mask] <- "FSGI"

# Failure to Stop and Render Aid
mask <- update_charge("F\\.?S\\.?R\\.?A")
DF_charges$l[mask] <- "FSRA"

mask <- update_charge(" AID")
DF_charges$l[mask] <- "FSRA"

mask <- update_charge("HIT AND RUN")
DF_charges$l[mask] <- "FSRA"

mask <- update_charge("LEAVING T?H?E? ?SCENE")
DF_charges$l[mask] <- "FSRA"

# Registration issue
mask <- update_charge("REGIS")
DF_charges$m[mask] <- "Registration Issue"

mask <- update_charge("EXPIRED REG")
DF_charges$m[mask] <- "Registration Issue"

mask <- update_charge("EXP?\\.? REG")
DF_charges$m[mask] <- "Registration Issue"

mask <- update_charge("NO REG\\.?$")
DF_charges$m[mask] <- "Registration Issue"

mask <- update_charge("EXP/NO REG")
DF_charges$m[mask] <- "Registration Issue"

# Unsafe Backing
mask <- update_charge("BACK")
DF_charges$n[mask] <- "Unsafe Backing"

# Follow Too Close
mask <- update_charge("TOO? CLOSS?[LE]E?Y?")
DF_charges$o[mask] <- "Follow Too Close"

# Stop Sign
mask <- update_charge("STOP SIGN")
DF_charges$p[mask] <- "Stop Sign"

mask <- update_charge("RAN STOP")
DF_charges$p[mask] <- "Stop Sign"

# Unsafe Turn
mask <- update_charge("TURN")
DF_charges$q[mask] <- "Unsafe Turn"

# Inspection
mask <- update_charge("INSP")
DF_charges$r[mask] <- "Inspection"

mask <- update_charge("MVI")
DF_charges$r[mask] <- "Inspection"

mask <- update_charge("EXP\\.?I?R?E?D? INS")
DF_charges$r[mask] <- "Inspection"

# License Plate
mask <- update_charge("MVR")
DF_charges$s[mask] <- "License Plate"

mask <- update_charge(" TAG")
DF_charges$s[mask] <- "License Plate"

# Unsafe Passing
mask <- update_charge("PASS ")
DF_charges$t[mask] <- "Unsafe Passing"

mask <- update_charge("PASSING")
DF_charges$t[mask] <- "Unsafe Passing"

mask <- update_charge("PASSED")
DF_charges$t[mask] <- "Unsafe Passing"

mask <- update_charge("DRIVING ON SHOULDER")
DF_charges$t[mask] <- "Unsafe Passing"

# Other Lane violations
mask <- update_charge("DESIGNATED LANE")
DF_charges$u[mask] <- "Other Lane Violations"

mask <- update_charge("ONE-HALF")
DF_charges$u[mask] <- "Other Lane Violations"

mask <- update_charge(" HALF ")
DF_charges$u[mask] <- "Other Lane Violations"

mask <- update_charge(" 1/2 ")
DF_charges$u[mask] <- "Other Lane Violations"

mask <- update_charge("SHOULDER")
DF_charges$u[mask] <- "Other Lane Violations"

mask <- update_charge("CENTER LANE")
DF_charges$u[mask] <- "Other Lane Violations"

# Child Seat
mask <- update_charge("CHILD")
DF_charges$v[mask] <- "Child Seat"

mask <- update_charge("CAR SEAT")
DF_charges$v[mask] <- "Child Seat"

# Texting etc
mask <- update_charge("TEXT")
DF_charges$w[mask] <- "Texting/Phone"

mask <- update_charge("PHONE")
DF_charges$w[mask] <- "Texting/Phone"

mask <- update_charge(" ELEC")
DF_charges$w[mask] <- "Texting/Phone"

# Unsafe start
mask <- update_charge("START")
DF_charges$x[mask] <- "Unsafe Start"

# Unsafe Load
mask <- update_charge("LOAD")
DF_charges$y[mask] <- "Unsafe Load"

mask <- update_charge("FAIL TO SECURE")
DF_charges$y[mask] <- "Unsafe Load"

# no seatbelt
mask <- update_charge("BELT")
DF_charges$z[mask] <- "No Seatbelt"

mask <- update_charge("RESTRAIN")
DF_charges$z[mask] <- "No Seatbelt"

# Evading
mask <- update_charge("EVAD")
DF_charges$A[mask] <- "Evading"

# Evasive Action
mask <- update_charge("EVASS?IVE")
DF_charges$B[mask] <- "Faulty Evasive Action"

# Address Change
mask <- update_charge("ADD")
DF_charges$C[mask] <- "Address Change"

mask <- update_charge("ADRESS")
DF_charges$C[mask] <- "Address Change"

# Wrong Way
mask <- update_charge("WRONG WAY")
DF_charges$D[mask] <- "Wrong Way"

mask <- update_charge("WRONG SIDE")
DF_charges$D[mask] <- "Wrong Way"

# Impeding Traffic
mask <- update_charge("IMPED")
DF_charges$E[mask] <- "Impeding Traffic"

# Inattention
mask <- update_charge("DISTRACT")
DF_charges$F[mask] <- "Distraction"

mask <- update_charge("INATT")
DF_charges$F[mask] <- "Distraction"

# Defective
mask <- update_charge("DEFECT")
DF_charges$G[mask] <- "Defective Vehicle"

mask <- update_charge("BALD")
DF_charges$G[mask] <- "Defective Vehicle"

mask <- update_charge("CRACK")
DF_charges$G[mask] <- "Defective Vehicle"

# Unauthorized Use
mask <- update_charge("UUMV")
DF_charges$H[mask] <- "Unauthorized Use"

mask <- update_charge("UNAUTHORIZED USE")
DF_charges$H[mask] <- "Unauthorized Use"

# Drug possession
mask <- update_charge("PARAPH")
DF_charges$I[mask] <- "Drug Possession"

mask <- update_charge("SUBST")
DF_charges$I[mask] <- "Drug Possession"

mask <- update_charge("MARIJ")
DF_charges$I[mask] <- "Drug Possession"

mask <- update_charge("^P\\.?O\\.?M")
DF_charges$I[mask] <- "Drug Possession"

mask <- update_charge("PCS")
DF_charges$I[mask] <- "Drug Possession"

mask <- update_charge("COCAINE")
DF_charges$I[mask] <- "Drug Possession"

mask <- update_charge("DANGEROUS DRUG")
DF_charges$I[mask] <- "Drug Possession"

# Open Container
mask <- update_charge("OPEN CON")
DF_charges$J[mask] <- "Open Container"

mask <- update_charge("OPEN ALCOH")
DF_charges$J[mask] <- "Open Container"

mask <- update_charge("POSS.* ALCOHOL")
DF_charges$J[mask] <- "Open Container"

# Reckless
mask <- update_charge("RECK")
DF_charges$K[mask] <- "Reckless"

mask <- update_charge("CARELESS")
DF_charges$K[mask] <- "Reckless"

# Stationary Object
mask <- update_charge("STATIONA")
DF_charges$L[mask] <- "Stationary Object"

mask <- update_charge("FIXED")
DF_charges$L[mask] <- "Stationary Object"

# Fail to Control
mask <- update_charge("FAILE?D? TO CONTROL$")
DF_charges$M[mask] <- "Fail to Control"

mask <- update_charge("FAILURE TO CONTROL$")
DF_charges$M[mask] <- "Fail to Control"

mask <- update_charge("CONTROL MOTOR")
DF_charges$M[mask] <- "Fail to Control"

# Fail to Stop properly
mask <- update_charge("STOP A?T? ?DESIG")
DF_charges$N[mask] <- "Improper Stop"

mask <- update_charge("STOP A?T? ?PROPER")
DF_charges$N[mask] <- "Improper Stop"

# Blocking Traffic
mask <- update_charge("BLOCK")
DF_charges$O[mask] <- "Blocking Traffic"

mask <- update_charge("PARK[EI][DN]G? [OI]N ROADWAY")
DF_charges$O[mask] <- "Blocking Traffic"

mask <- update_charge("PARK[EI][DN]G? IN A? ?TRAFFIC")
DF_charges$O[mask] <- "Blocking Traffic"

mask <- update_charge("PARK[EI][DN]G? IN A? ?MOVING")
DF_charges$O[mask] <- "Blocking Traffic"

mask <- update_charge("STOPP?E?D? IN A? ?MOVING")
DF_charges$O[mask] <- "Blocking Traffic"

# Misc Criminal charge
mask <- update_charge("CRIMINAL")
DF_charges$P[mask] <- "Misc Criminal"

mask <- update_charge("ASSAULT")
DF_charges$P[mask] <- "Misc Criminal"

mask <- update_charge("THEFT")
DF_charges$P[mask] <- "Misc Criminal"

mask <- update_charge("ROBBERY")
DF_charges$P[mask] <- "Misc Criminal"

mask <- update_charge("RESISTING ARREST")
DF_charges$P[mask] <- "Misc Criminal"

mask <- update_charge("MANSLAUGHTER")
DF_charges$P[mask] <- "Misc Criminal"

mask <- update_charge("MURDER")
DF_charges$P[mask] <- "Misc Criminal"

mask <- update_charge("WEAPON")
DF_charges$P[mask] <- "Misc Criminal"

mask <- update_charge("FIREARM")
DF_charges$P[mask] <- "Misc Criminal"

mask <- update_charge("UCW")
DF_charges$P[mask] <- "Misc Criminal"

mask <- update_charge("WARRANT")
DF_charges$P[mask] <- "Misc Criminal"

# Safety Zone
mask <- update_charge("SAFETY ZONE")
DF_charges$Q[mask] <- "Safety Zone"

# Parked Facing Traffic
mask <- update_charge("FACING")
DF_charges$R[mask] <- "Park Facing Traf"

# Failure to Report
mask <- update_charge("REPORT ACC")
DF_charges$S[mask] <- "Failure to Report"

mask <- update_charge("COMPLY")
DF_charges$S[mask] <- "Failure to Report"

mask <- update_charge("NOTIFY OWNER")
DF_charges$S[mask] <- "Failure to Report"

mask <- update_charge("DUTY U?P?ON")
DF_charges$S[mask] <- "Failure to Report"

# Racing
mask <- update_charge("RACING")
DF_charges$T[mask] <- "Racing"

# No Adult
mask <- update_charge("CODE B")
DF_charges$U[mask] <- "No Adult"

mask <- update_charge("RESTRICTION B")
DF_charges$U[mask] <- "No Adult"

mask <- update_charge("LOFS 21")
DF_charges$U[mask] <- "No Adult"

# Pedestrian error
mask <- update_charge("PEDESTRIAN ENTER")
DF_charges$V[mask] <- "Pedestrian Error"

mask <- update_charge("PEDESTRIAN CROSSED")
DF_charges$V[mask] <- "Pedestrian Error"

mask <- update_charge("PEDESTRIAN IN ROAD")
DF_charges$V[mask] <- "Pedestrian Error"

mask <- update_charge("JAY")
DF_charges$V[mask] <- "Pedestrian Error"

# Pedestrian error
mask <- update_charge("WITHOUT H?E?A?D? ?LIGHTS")
DF_charges$W[mask] <- "Lights Off"

mask <- update_charge("W/O H?E?A?D? ?LIGHTS")
DF_charges$W[mask] <- "Lights Off"

mask <- update_charge("NO H?E?A?D? ?LIGHTS")
DF_charges$W[mask] <- "Lights Off"

# Toll Violation
mask <- update_charge(" TOLL")
DF_charges$X[mask] <- "Toll Violation"

# Obey Officer
mask <- update_charge("OFFICER")
DF_charges$Y[mask] <- "Obey Officer"

mask <- update_charge("SCHOOL CROSS")
DF_charges$Y[mask] <- "Obey Officer"



# Collapse a,... columns into Simple_Charge

DF_Charges_temp <- 
  DF_charges %>% 
  tidyr::unite("Simple_Charge", a:Y, na.rm = TRUE, remove = FALSE, sep=", ")

#   Look at what is not categorized
ChargesTemp <- DF_Charges_temp[DF_Charges_temp$Simple_Charge=="",] %>% group_by(Charge) %>% tally() %>% arrange(-n)


#   And when done, save the final

foo <- 
  DF_charges %>% 
  tidyr::unite("Simple_Charge", a:Y, na.rm = TRUE, remove = FALSE, sep=", ") %>% 
  select(-a:-Y)

DF_charges <- foo

DF_charges$Simple_Charge <- str_replace(DF_charges$Simple_Charge,"^, ", "")
DF_charges$Simple_Charge <- str_replace(DF_charges$Simple_Charge,"^ $", "")
DF_charges$Simple_Charge <- str_replace(DF_charges$Simple_Charge,"^$", "UNK")

```


```{r now finish putting datasets together}

# need to rearrange charges so that they show up as a list for each
# crash.

DF_charges <- DF_charges %>% 
  group_by(Crash_ID) %>% 
  arrange(Simple_Charge) %>% 
  summarise(Charges = paste(Simple_Charge, collapse=", "))

# combine the two

DF <- left_join(DF_collisions, DF_charges, by="Crash_ID")

df <- DF %>% select(Crash_Fatal_Fl, Crash_Date,	Crash_Time, Rpt_City_ID, Rpt_Block_Num, Rpt_Street_Pfx,	Rpt_Street_Name, Rpt_Street_Sfx, At_Intrsct_Fl, Latitude, Longitude, Street_Name, Street_Nbr, Tot_Injry_Cnt, Death_Cnt, Crash_Sev_ID, Crash_Speed_Limit, Harm_Evnt_ID, Road_Cls_ID, Intrsct_Relat_ID, Nbr_Of_Lane, Charges  )

df <- left_join(df, City, by="Rpt_City_ID")
df <- left_join(df, Severity, by="Crash_Sev_ID")
df <- left_join(df, Harm, by="Harm_Evnt_ID")
df <- left_join(df, RoadType, by="Road_Cls_ID")
df <- left_join(df, Intersection, by="Intrsct_Relat_ID")



```


Do some quality control checks


```{r Look for nulls}

# Look for NA's in the various fields

df %>%
    map_df(function(x) sum(is.na(x))) %>%
    gather(feature, num_nulls) %>%
    print(n = 100)

```

```{r Look at Crash_Fatal_Fl}

sort(unique(df$Crash_Fatal_Fl))

```

```{r Look at Crash_Date}

sort(unique(df$Crash_Date))

range(df$Crash_Date)
range(df$Crash_Time)

```


```{r Look at Crash_Time}

sort(unique(df$Crash_Time))

```

```{r Look at City ID}

sort(unique(df$Rpt_City_ID))

```

```{r Look at Rpt_Blk_Num}

sort(unique(df$Rpt_Block_Num))

df$Rpt_Block_Num[grepl("\\D+", df$Rpt_Block_Num)]

#   Remove non-numeric stuff

df$Rpt_Block_Num <- str_replace(df$Rpt_Block_Num, " BLOCK","")
df$Rpt_Block_Num <- str_replace_all(df$Rpt_Block_Num, "O","0")
df$Rpt_Block_Num <- str_replace_all(df$Rpt_Block_Num, "^\\d+-","")
df$Rpt_Block_Num <- str_replace_all(df$Rpt_Block_Num, "PP","00")
df$Rpt_Block_Num <- str_replace_all(df$Rpt_Block_Num, " 1/2","")
df$Rpt_Block_Num <- str_replace(df$Rpt_Block_Num, "\\D+","")
df$Rpt_Block_Num <- str_replace(df$Rpt_Block_Num, "0RE","")
df$Rpt_Block_Num <- str_replace(df$Rpt_Block_Num, "0NE","")
df$Rpt_Block_Num <- str_replace(df$Rpt_Block_Num, "#","")

sort(unique(df$Rpt_Block_Num))
```

```{r Look at Rpr_Street_Pfx}

sort(unique(df$Rpt_Street_Pfx))

df$Rpt_Street_Pfx <- str_replace(df$Rpt_Street_Pfx, "UNK", "")
```
```{r Look at Rpt_Street_Sfx}

sort(unique(df$Rpt_Street_Sfx))

```

```{r Look at Latitude and Longitude}

df$Latitude[grepl("[^0-9\\-\\. ]", df$Latitude)]
df$Longitude[grepl("[^-][^0-9\\. ]", df$Longitude)]
df$Latitude <- as.numeric(df$Latitude)
df$Longitude <- as.numeric(df$Longitude)
range(df$Latitude, na.rm=TRUE)
range(df$Longitude, na.rm = TRUE)

```
```{r Look at Street_Nbr}

sort(unique(df$Street_Nbr))
df$Street_Nbr[grepl("\\D+", df$Street_Nbr)]
df$Street_Nbr <- str_replace(df$Street_Nbr, "\\D+","")

```


```{r Look at Tot_Injry_Cnt}

sort(unique(df$Tot_Injry_Cnt))

```

```{r Look at Death_Cnt}

sort(unique(df$Death_Cnt))

```

```{r Look at Severity}

sort(unique(df$Severity))

df %>% group_by(Severity) %>% 
  summarise(count=n()) %>%
  arrange(count)

```


```{r Look at Harmed}

sort(unique(df$Harmed))

df %>% group_by(Harmed) %>% 
  summarise(count=n()) %>%
  arrange(count)

```


```{r Look at RoadType}

sort(unique(df$RoadType))
```



```{r Look at Intersection}

sort(unique(df$Intersection))
```


```{r Look at Number of Lanes}

sort(unique(df$Nbr_Of_Lane))
summary(df$Nbr_Of_Lane)
```

```{r Look at Rpt_Street_Name}

sort(unique(df$Rpt_Street_Name))
```

```{r Look at Street_Name}

sort(unique((df$Street_Name)))

```


```{r save dataset}

saveRDS(df, "~/Dropbox/Rprojects/Collisions/Collisions_2010_2018.RDS")

```


#     Add neighborhood and zipcode designations to records

#```{r Add neighborhood and zipcode}
####   mask out rows with bad coordinates
maskcoord <- !(is.na(df$Latitude) | is.na(df$Longitude))
#  Create a temporary sf data frame for doing the intersects
# set longitudes as the first column and latitudes as the second
dat <- data.frame(Longitude=df$Longitude[maskcoord], Latitude=df$Latitude[maskcoord], stringsAsFactors = FALSE)

dat <- st_as_sf(dat, coords=c("Longitude", "Latitude"), crs=googlecrs, agr = "identity")
# read in neighborhood and zipcode polygons
neighborhoods <- readRDS("~/Dropbox/CrimeStats/NeighborhoodPolys.rds")
zipcodes <- readRDS("~/Dropbox/CrimeStats/ZipCodes.rds")

# st_intersects to see if points are inside poly
#  This is a matrix of T/F values, npoints x npolys in size
#  if a point is in multiple polys it will show up here
a <- st_intersects(dat, neighborhoods, sparse = FALSE)
max(rowSums(a)) # this is the maximum number of polys for some point
hist(rowSums(a))# just curious how they are distributed.
#   collect all the neighborhoods a point falls into, separate by
#   commas, and put into the Nbhd field
diddle <- function (i) {paste(neighborhoods$Name[i], collapse=', ')}
df$Nbhd[maskcoord] <-  apply(a,1,diddle)  #  work on "a" row-wise

######################################
#########   add zipcode
######################################
#   find points in polygons
#   since zipcodes don't overlap, let's just grab the array index
#   instead of creating a huge matrix
a <- st_intersects(dat, zipcodes, sparse = TRUE)

#   Append the ZIP field to the data frame
df$Zip_Code[maskcoord] <- zipcodes$Zip_Code[unlist(a)]


#```


```{r save result out to a file}

saveRDS(df, file="~/Dropbox/Rprojects/Collisions/Collisions_2010_2017nozip.rds")
```

